---
title: "MassWateR analyses"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MassWateR analyses}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, message = F, warning = F}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  warning = T, 
  message = F
)
library(ggplot2)
```

The analysis functions in MassWateR can be used to evaluate trends, summaries, and maps once the required data are successfully imported into R (see the [data input and checks vignette](https://massbays-tech.github.io/MassWateR/articles/MassWateR.html) for an overview).  The results data file with the monitoring data is required.  The data quality objectives file for accuracy is also required to automatically determine plot axis scaling as arithmetic (linear) or logarithmic and to fill results data that are below detection or above quantitation limits. The site metadata file is used to create maps. The example data included with the package are imported here to demonstrate how to use the analysis functions:

```{r, message = T}
library(MassWateR)

# import results data
respth <- system.file("extdata/ExampleResults.xlsx", package = "MassWateR")
resdat <- readMWRresults(respth)

# import accuracy data
accpth <- system.file("extdata/ExampleDQOAccuracy.xlsx", package = "MassWateR")
accdat <- readMWRacc(accpth)

# import site metadata
sitpth <- system.file("extdata/ExampleSites.xlsx", package = "MassWateR")
sitdat <- readMWRsites(sitpth)
```

## Analyze seasonal trends

Seasonal trends can be evaluated using the `anlzMWRseason()` function.  This function summarizes results for a single parameter using boxplots or barplots with seasonal groups assigned to months or weeks of the year.  Boxplots or barplots can also include jittered points of the observations on top or only the jittered points can be shown.  Threshold lines are also shown on the plot to determine if results are near thresholds of interest.  The data that are plotted can be filtered by sites, dates, result attributes, or location groups (from the site metadata file).

The `anlzMWRseason()` function uses the results data as a file path or data frame as input. Below, the results for dissolved oxygen (`param = "DO"`) are grouped by month using `group = "month"` and as boxplots using `type = "box"`.  A description of the summary statistics shown by a boxplot is provided in the [outlier checks](https://massbays-tech.github.io/MassWateR/articles/outlierchecks.html#analyzing-outliers) vignette.  The relevant threshold lines for freshwater environments are shown using `thresh = "fresh"`.  

```{r, fig.height = 4, fig.width = 8}
anlzMWRseason(res = resdat, param = "DO", acc = accdat, thresh = "fresh", group = "month", type = "box")
```

Jittered points over the boxplots can be shown by setting `type = "jitterbox"`.

```{r, fig.height = 4, fig.width = 8}
anlzMWRseason(res = resdat, param = "DO", acc = accdat, thresh = "fresh", group = "month", type = "jitterbox")
```

The same data can also be grouped by week using `group = "week"`.  The week of the year is shown on the plot as an integer. Note that there can be no common month/day indicating the start of the week between years and an integer is the only way to compare summaries if the results data span multiple years. 

```{r, fig.height = 4, fig.width = 8}
anlzMWRseason(res = resdat, param = "DO", acc = accdat, thresh = "fresh", group = "week", type = "box")
```

Barplots can be shown using `type = "bar"`.  The barplots show the average estimate for the results defined by the `group` argument.  

```{r, fig.height = 4, fig.width = 8}
anlzMWRseason(res = resdat, param = "DO", acc = accdat, thresh = "fresh", group = "month", type = "bar")
```

Confidence intervals at 95% for the barplots can be shown by setting `confint = TRUE`.

```{r, fig.height = 4, fig.width = 8}
anlzMWRseason(res = resdat, param = "DO", acc = accdat, thresh = "fresh", group = "month", type = "bar", confint = TRUE)
```

Jittered points over the barplots can be shown by setting `type = "jitterbar"`.

```{r, fig.height = 4, fig.width = 8}
anlzMWRseason(res = resdat, param = "DO", acc = accdat, thresh = "fresh", group = "month", type = "jitterbar")
```

Setting `type = "jitter"` will show only the jittered points.

```{r, fig.height = 4, fig.width = 8}
anlzMWRseason(res = resdat, param = "DO", acc = accdat, thresh = "fresh", group = "month", type = "jitter")
```

The threshold lines shown on each plot are determined from the `thresholdMWR` file included with MassWateR.  These threshold lines describe relevant state standards or typical ranges in Massachusetts, if applicable for a parameter. Thresholds can be plotted for freshwater or marine environments using the `thresh` argument as `thresh = "fresh"` or `thresh = "marine"`, respectively.  Two threshold lines are typically shown, whereas some parameters may only have one threshold line. Other parameters have no information in the `thresholdMWR` file and lines will not be plotted. Threshold lines can be suppressed by setting `thresh = "none"`. Below, the marine thresholds are shown for dissolved oxygen using the barplot option grouped by month. 

```{r, fig.height = 4, fig.width = 8}
anlzMWRseason(res = resdat, param = "DO", acc = accdat, thresh = "marine", group = "month", type = "bar")
```

The y-axis scaling as arithmetic (linear) or logarithmic can be set with the `yscl` argument.  If `yscl = "auto"` (default), the scaling is determined automatically from the data quality objective file for accuracy, i.e., parameters with "log" in any of the columns are plotted on log10-scale, otherwise arithmetic. Setting `yscl = "linear"` or `yscl = "log"` will set the axis as linear or log10-scale, respectively, regardless of the information in the data quality objective file for accuracy. Below, the axis for E. Coli is plotted on the log-10 scale automatically.  The y-axis scaling does not need to specified explicitly in the function call because the default setting is `yscl = "auto"`.

```{r, fig.height = 4, fig.width = 8}
anlzMWRseason(res = resdat, param = "E.coli", acc = accdat, thresh = "fresh", group = "month", type = "bar")
```

To force the y-axis as linear for E. Coli, `yscl = "linear"` must be used. Note that the default linear scaling for dissolved oxygen above is determined automatically. The means and confidence intervals will also vary between arithmetic and log-scaling if `type = "bar"`.

```{r, fig.height = 4, fig.width = 8}
anlzMWRseason(res = resdat, param = "E.coli", acc = accdat, thresh = "fresh", group = "month", type = "bar", yscl = "linear")
```

Results can also be filtered by dates using the `dtrng` argument. The date format must be `YYYY-MM-DD` and include two entries. 

```{r, fig.height = 4, fig.width = 8}
anlzMWRseason(res = resdat, param = "E.coli", acc = accdat, thresh = "fresh", group = "month", type = "bar", dtrng = c("2021-05-01", "2021-07-31"))
```

Finally, the results can be filtered by sites, result attributes, and location groups using the `site`, `resultatt`, and `locgroup` arguments. These options will filter the results data based on values in the `Monitoring Location ID` and `Result Attribute` columns in the results file and by the `Location Group` column in the site metadata file, which must be provided to the `sit` argument if `locgroup` is used.  

```{r, fig.height = 4, fig.width = 8}
anlzMWRseason(res = resdat, param = "E.coli", acc = accdat, sit = sitdat, thresh = "fresh", group = "month", site = "ABT-077", resultatt = "Dry", locgroup = "Lower Assabet")
```

## Analyze trends by date

Trends by date for selected parameters can be evaluated using the `anlzMWRdate()` function.  The plot shows results continuously over time as line plots, with the lines separated by site, combined across sites, or combined across location groups as an average with 95% confidence intervals.  For location groups, the site metadata file must be supplied.  Many of the options described for the `anlzMWRseason()` function also apply and will not be repeated here. As before, the required data are a results file and data quality objective file for accuracy.  

The default plot for `anlzMWRdate()` will show separate lines for each site, y-axis scaling determined automatically from the data quality objectives file, and site labels at the end of each line.  All sites in the results file will be plotted, which can be difficult to visualize. Selecting specific sites with the `site` argument is recommended. 

```{r, fig.height = 4, fig.width = 8}
anlzMWRdate(res = resdat, param = "DO", acc = accdat, thresh = "fresh", group = "site", site = c("ABT-026", "ABT-077"))
```

A color legend for the sites can be included by setting `colleg = TRUE`.

```{r, fig.height = 4, fig.width = 8}
anlzMWRdate(res = resdat, param = "DO", acc = accdat, thresh = "fresh", group = "site", site = c("ABT-026", "ABT-077"),
            colleg = T)
```

The results averaged across the selected sites can be plotted using `group = "all"`.

```{r, fig.height = 4, fig.width = 8}
anlzMWRdate(res = resdat, param = "DO", acc = accdat, thresh = "fresh", group = "all", site = c("ABT-026", "ABT-077"))
```

The results averaged across all sites in the results file can be plotted by omitting the `site` argument.

```{r, fig.height = 4, fig.width = 8}
anlzMWRdate(res = resdat, param = "DO", acc = accdat, thresh = "fresh", group = "all")
```

Confidence intervals at 95% for the points can be shown when `group = "all"` or `"locgroup"` by setting `confint = TRUE`.

```{r, fig.height = 4, fig.width = 8}
anlzMWRdate(res = resdat, param = "DO", acc = accdat, thresh = "fresh", group = "all", confint = TRUE)
```

Sites can be averaged across location groups using `group = "locgroup"`.  This option requires the site metadata file passed to the `sit` argument and appropriate values for the location groups passed to the `locgroup` argument that match values in the `Location Group` column of the site metadata file. 

```{r, fig.height = 4, fig.width = 8}
anlzMWRdate(res = resdat, param = 'DO', acc = accdat, sit = sitdat, group = 'locgroup', 
      thresh = 'fresh', locgroup = c('Lower Assabet', 'Upper Assabet'))
```

All location groups can be used if the `locgroup` argument is omitted. 

```{r, fig.height = 4, fig.width = 8}
anlzMWRdate(res = resdat, param = 'DO', acc = accdat, sit = sitdat, group = 'locgroup', 
      thresh = 'fresh')
```

## Analyze trends by site

Trends by site for selected parameters can be evaluated using the `anlzMWRsite()` function.  This function summarizes results for a single parameter using boxplots or barplots separately for each site on the x-axis. Boxplots or barplots can also include jittered points of the observations on top or only the jittered points can be shown. Many of the options described for the `anlzMWRseason()` function also apply and will not be repeated here. As before, the required data are a results file and data quality objective file for accuracy.  

The default plot for `anlzMWRsite()` will show boxplots of results for each site and y-axis scaling determined automatically from the data quality objectives file. 

```{r, fig.height = 4, fig.width = 8}
anlzMWRsite(res = resdat, param = "DO", acc = accdat, thresh = "fresh", type = "box")
```

Jittered points over the boxplots can be shown by setting `type = "jitterbox"`.

```{r, fig.height = 4, fig.width = 8}
anlzMWRsite(res = resdat, param = "DO", acc = accdat, thresh = "fresh", type = "jitterbox")
```

Results as barplots can be shown using `type = "bar"`.

```{r, fig.height = 4, fig.width = 8}
anlzMWRsite(res = resdat, param = "DO", acc = accdat, thresh = "fresh", type = "bar")
```

Confidence intervals at 95% for the barplots can be shown by setting `confint = TRUE`.

```{r, fig.height = 4, fig.width = 8}
anlzMWRsite(res = resdat, param = "DO", acc = accdat, thresh = "fresh", type = "bar", confint = TRUE)
```

Jittered points over the barplots can be shown by setting `type = "jitterbar"`.

```{r, fig.height = 4, fig.width = 8}
anlzMWRsite(res = resdat, param = "DO", acc = accdat, thresh = "fresh", type = "jitterbar")
```

Setting `type = "jitter"` will show only the jittered points.

```{r, fig.height = 4, fig.width = 8}
anlzMWRsite(res = resdat, param = "DO", acc = accdat, thresh = "fresh", type = "jitter")
```

Results can be grouped by entries in the `Result Atribute` column using `byresultatt = TRUE`. For example, sites with E. coli samples can be grouped by `"Dry"` or `"Wet"` conditions.  Filtering by sites first using the `site` argument is advised to reduce the amount of data that are plotted. The grouping can be filtered further by passing appropriate values in the `Result Attribute` column to the `resultatt` argument. Note that specifying result attributes with `resultatt` and setting `byresultatt = FALSE` will filter the plot data by the result attributes but will not plot the results separately. 

```{r, fig.height = 4, fig.width = 8}
anlzMWRsite(res = resdat, param = "E.coli", acc = accdat, thresh = "fresh", type = "box", 
                        site = c("ABT-077", "ABT-162", "CND-009", "CND-110", "HBS-016", "HBS-031"),
                        resultatt = c('Dry', 'Wet'), byresultatt = TRUE)
```

## Analyze results with maps

The `anlzMWRmap()` function can be used to create a map of summarized results for a selected parameter at each monitoring site.  By default, all dates for the parameter are averaged. Options to filter by site, date range, and result attribute are provided.  Only sites with spatial information in the site metadata file are plotted and a warning is returned for those that do not have this information. The site labels are also plotted next to each point. 

A map of dissolved oxygen averages across all dates can be created as follows. The required inputs are the results data, the data quality objective file for accuracy, and the site metadata file with the site locations. The warning message indicates some sites in the results file do not have spatial data in the site metadata file.

```{r, fig.height = 7, fig.width = 6}
anlzMWRmap(res = resdat, param = "DO", acc = accdat, sit = sitdat, addwater = "nhd")
```

By default, lines and polygons of natural water bodies defined using the National Hydrography Dataset (NHD) are plotted using the `addwater = "nhd"` argument.  These are are included as data files with the package as `pondsMWR`, `riversMWR`, and `streamsMWR`. The level of detail for the water bodies is defined by the `dLevel` argument as `"low"`, `"medium"` (default), or `"high"`.  The detail level only applies for the NHD data. 

Water body detail as low: 

```{r, fig.height = 7, fig.width = 6}
anlzMWRmap(res = resdat, param = "DO", acc = accdat, sit = sitdat, addwater = "nhd", dLevel = "low")
```

Water body detail as high: 

```{r, fig.height = 7, fig.width = 6}
anlzMWRmap(res = resdat, param = "DO", acc = accdat, sit = sitdat, addwater = "nhd", dLevel = "high")
```

Using `addwater = "osm"` will include lines and polygons of natural water bodies from the OpenStreetMap ([OSM](https://www.openstreetmap.org/)) project, downloaded using the [osmdata](https://docs.ropensci.org/osmdata) package. The former may be preferred for more accurate display of water bodies in Massachusetts, whereas the latter option will also include water bodies outside of the state. Use `addwater = NULL` to suppress.

Using `addwater = "osm"` (warnings suppressed):

```{r, fig.height = 7, fig.width = 6}
anlzMWRmap(res = resdat, param = "DO", acc = accdat, sit = sitdat, addwater = "osm", warn = F)
```

A north arrow, scale bar, and labels are also plotted.  These can be suppressed by setting the appropriate arguments to `NULL`.

```{r, fig.height = 7, fig.width = 6}
anlzMWRmap(res = resdat, param = "DO", acc = accdat, sit = sitdat, northloc = NULL, scaleloc = NULL, labsize = NULL, warn = F)
```

Locations of the north arrow and scale bar can be changed with the same arguments by specifying `"tl"`, `"tr"`, `"bl"`, or `"br"` for top-left, top-right, bottom-left, or bottom-right. 

```{r, fig.height = 7, fig.width = 6}
anlzMWRmap(res = resdat, param = "DO", acc = accdat, sit = sitdat, northloc = "br", scaleloc = "tl", warn = F)
```

The latitude and longitude text on the plot axes can be suppressed using `latlon = FALSE`.

```{r, fig.height = 7, fig.width = 6}
anlzMWRmap(res = resdat, param = "DO", acc = accdat, sit = sitdat, latlon = F, warn = F)
```

The color palette for the average values can also be changed.  Any palette from [RColorBrewer](https://r-graph-gallery.com/38-rcolorbrewers-palettes.html) can be used. 

```{r, fig.height = 7, fig.width = 6}
anlzMWRmap(res = resdat, param = "DO", acc = accdat, sit = sitdat, palcol = "Spectral", warn = F)
```

The buffered distance around the points can be increased uzing the `buffdist` argument. 

```{r, fig.height = 7, fig.width = 7}
anlzMWRmap(res = resdat, param = "DO", acc = accdat, sit = sitdat, buffdist = 0.3, warn = F)
```

A base map can be included as well using the `maptype` argument. Options include any stamen map as `"terrain"`, `"terrain-background"` (default), `"terrain-labels"`, `"terrain-lines"`, `"toner"`, `"toner-2010"`, `"toner-2011"`, `"toner-background"`, `"toner-hybrid"`, `"toner-labels"`, `"toner-lines"`, `"toner-lite"`, or `"watercolor"`.  The `zoom` argument can be helpful when using a basemap. The default `zoom` argument is set to 11 and decreasing the number will download a base map with lower resolution.  This can decrease map processing times for large areas.

A `"terrain"` base map: 

```{r, fig.height = 7, fig.width = 6}
anlzMWRmap(res = resdat, param = "DO", acc = accdat, sit = sitdat, maptype = "terrain", warn = F, addwater = NULL)
```

A `"terrain-lines"` base map: 

```{r, fig.height = 7, fig.width = 6}
anlzMWRmap(res = resdat, param = "DO", acc = accdat, sit = sitdat, maptype = "terrain-lines", warn = F, addwater = NULL)
```

A `"toner"` base map:

```{r, fig.height = 7, fig.width = 6}
anlzMWRmap(res = resdat, param = "DO", acc = accdat, sit = sitdat, maptype = "toner", warn = F, addwater = NULL)
```

A `"toner-labels"` base map:

```{r, fig.height = 7, fig.width = 6}
anlzMWRmap(res = resdat, param = "DO", acc = accdat, sit = sitdat, maptype = "toner-labels", warn = F, addwater = NULL)
```

A `"watercolor"` base map:

```{r, fig.height = 7, fig.width = 6}
anlzMWRmap(res = resdat, param = "DO", acc = accdat, sit = sitdat, maptype = "watercolor", warn = F, addwater = NULL)
```

A map with no base map or water bodies:

```{r, fig.height = 7, fig.width = 6}
anlzMWRmap(res = resdat, param = "DO", acc = accdat, sit = sitdat, maptype = NULL, warn = F, addwater = NULL)
```
