---
title: "MassWateR quality control functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MassWateR quality control functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  warning = F, 
  message = F
)
```

The quality control functions in MassWateR can be used once the required data are successfully imported into R (see the [previous vignette](https://massbays-tech.github.io/MassWateR/articles/datainput.html) for an overview).  The required data includes the results data file that describe the monitoring data and the data quality objective files for completeness and accuracy.  The example data files included with the package are imported here to demonstrate how to use the quality control functions:

```{r, message = T}
library(MassWateR)

# import results data
respth <- system.file('extdata/ExampleResults_final.xlsx', package = 'MassWateR')
resdat <- read_results(respth)

# import data quality objectives for accuracy
dqoaccpth <- system.file('extdata/ExampleDQOAccuracy_final.xlsx', package = 'MassWateR')
dqoaccdat <- read_dqoaccuracy(dqoaccpth)

# import data quality objectives for completeness
dqocompth <- system.file('extdata/ExampleDQOCompleteness_final.xlsx', package = 'MassWateR')
dqocomdat <- read_dqocompleteness(dqocompth)
```

## Quality control for accuracy

## Quality control for completeness