---
title: "MassWateR quality control functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MassWateR quality control functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  warning = T, 
  message = F
)
```

The quality control functions in MassWateR can be used once the required data are successfully imported into R (see the [previous vignette](https://massbays-tech.github.io/MassWateR/articles/datainput.html) for an overview).  The required data includes the results data file that describe the monitoring data and the data quality objective files for accuracy, frequency, and completeness.  The example data files included with the package are imported here to demonstrate how to use the quality control functions:

```{r, message = T}
library(MassWateR)

# import results data
respth <- system.file('extdata/ExampleResults.xlsx', package = 'MassWateR')
resdat <- readMWRresults(respth)

# import data quality objectives for accuracy
accpth <- system.file('extdata/ExampleDQOAccuracy.xlsx', package = 'MassWateR')
accdat <- readMWRacc(accpth)

# import data quality objectives for frequency and completeness
frecompth <- system.file('extdata/ExampleDQOFrequencyCompleteness.xlsx', package = 'MassWateR')
frecomdat <- readMWRfrecom(frecompth)
```

## Quality control for accuracy

The quality control checks for accuracy assess several characteristics of the data in the results file by referencing appropriate values in the data quality objectives file for accuracy.  In short, the accuracy checks evaluate field blanks, lab blanks, field duplicates, lab duplicates, lab spikes, and instrument checks. The accuracy checks require results data (as in `resdat` above) and a data quality objectives file for accuracy (as in `accdat`).

The `tabMWRacc()` function is used to create tabular results for the accuracy checks for each parameter.  The function can be used with inputs as paths to the relevant files or as data frames returned by `readMWRresults()` and `readMWRacc()`.  For the former, the full suite of data checks can be evaluated with `runkchk = T` (default) or suppressed with `runchk = F`, as explained in the [data inputs and checks vignette](https://massbays-tech.github.io/MassWateR/articles/datainput.html).  In the latter case, downstream analyses may not work if data are formatted incorrectly.  Also note that accuracy is only evaluated on parameters that are shared between the results file and the data quality objectives accuracy file.  A warning is returned if there are parameters that do not match. The warnings can be suppressed by setting `warn = FALSE`. 

The `tabMWRacc()` function can return three types of tables as specified with the `type` argument: `"individual"`, `"summary"`, or `"percent"`.  The individual tables are specific to each type of accuracy check for each parameter (e.g., field blanks, lab blanks, etc.).  The summary table summarizes all accuracy checks by the number of checks and how many hit/misses are returned for each across all parameters.  The percent table is similar to the summary table, but showing only percentages with appropriate color-coding for hit/misses.   

For `type = "individual"`, the quality control tables for accuracy are retrieved by specifying the check with the `accchk` argument.  The `accchk` argument can be used to specify one of the following values to retrieve the relevant tables: `"Field Blanks"`, `"Lab Blanks"`, `"Field Duplicates"`, `"Lab Duplicates"`, `"Lab Spikes"`, or `"Instrument Checks (post sampling)"`.  Below shows how to retrieve each table using the data frames from above for the results and quality objectives file for accuracy. The warnings are suppressed with `warn = FALSE`.

```{r}
tabMWRacc(res = resdat, acc = accdat, type = 'individual', accchk = 'Field Blanks', warn = FALSE)
```
```{r}
tabMWRacc(res = resdat, acc = accdat, type = 'individual', accchk = 'Lab Blanks', warn = FALSE)
```
```{r}
tabMWRacc(res = resdat, acc = accdat, type = 'individual', accchk = 'Field Duplicates', warn = FALSE)
```
```{r}
tabMWRacc(res = resdat, acc = accdat, type = 'individual', accchk = 'Lab Duplicates', warn = FALSE)
```
```{r}
tabMWRacc(res = resdat, acc = accdat, type = 'individual', accchk = 'Lab Spikes', warn = FALSE)
```
```{r}
tabMWRacc(res = resdat, acc = accdat, type = 'individual', accchk = 'Instrument Checks (post sampling)', warn = FALSE)
```

For `type = "summary"`, the function summarizes all accuracy checks by counting the number of quality control checks, number of misses, and percent acceptance for each parameter.  The `accchk` argument can be used for one to any of all accuracy checks, defaulting to all if left blank.

```{r}
tabMWRacc(res = resdat, acc = accdat, type = 'summary', warn = FALSE)
```

For `type = "percent"`, the `tabMWRacc()` function... The `accchk` argument can be used for one to any of all accuracy checks, defaulting to all if left blank.

The `tabMWRacc()` function uses the `qcMWRacc()` function internally to create the table.  This function creates the raw summaries of accuracy from the input data.  Typically, `qcMWRacc()` is not used by itself, but it is explained here to demonstrate how the raw summaries can be obtained. 

Below, the `qcMWRacc()` function is executed with the data frames for the results and quality objectives file for accuracy.  As with `tabMWRacc`, the `accchk` argument can be used to return one to all of the checks.  The results are returned as a list, with each element of the list corresponding to a specific accuracy check.  Here, the lab duplicate checks are returned.  The warnings are also suppressed. 

```{r, warning = T}
qcMWRacc(res = resdat, acc = accdat, accchk = 'Lab Duplicates', warn = FALSE)
```

## Quality control for frequency 

The quality control checks for frequency are used to verify an appropriate number of quality control samples have been collected or analyzed for each parameter.  These are checks on the quantity of samples and not the values, as for the accuracy checks.  The frequency checks require results data (as in `resdat` above) and a data quality objectives file for frequency and completeness (as in `frecomdat`).

The `tabMWRfre()` function is used to create tabular results for the frequency checks for each parameter.  The function can be used with inputs as paths to the relevant files or as data frames returned by `readMWRresults()` and `readMWRfrecom()`.  For the former, the full suite of data checks can be evaluated with `runkchk = T` (default) or suppressed with `runchk = F`, as explained in the [data inputs and checks vignette](https://massbays-tech.github.io/MassWateR/articles/datainput.html).  In the latter case, downstream analyses may not work if data are formatted incorrectly.  Also note that completeness is only evaluated on parameters that are shared between the results file and the data quality objectives completeness file.  A warning is returned if there are parameters that do not match. The warnings can be suppressed by setting `warn = FALSE`. 

The quality control tables for frequency show the number of records that apply to a given check (e.g., Lab Blank, Field Blank, etc.) relative to the number of "regular" data records (e.g., field samples or measures) for each parameter.  A summary of all frequency checks for each parameter is provided if `type = "summary"`. The function is executed with the data frames for the results and quality objectives file for frequency. 

```{r}
tabMWRfre(res = resdat, frecom = frecomdat, type = 'summary', warn = F)
```

A color-coded table showing similar information as percentages for each parameter is provided if `type = "percent"`. Cells are shown as green or red if the required frequency checks are met as specified in the data quality objectives file.

```{r}
tabMWRfre(res = resdat, frecom = frecomdat, type = 'percent')
```

The `tabMWRfre()` function uses the `qcMWRfre()` function internally to create the table.  This function creates the raw summaries of frequency from the input data.  Typically, `qcMWRfre()` is not used by itself, but it is explained here to demonstrate how the raw summaries can be obtained. 

Below, the `qcMWRfre()` function is executed with the data frames for the results and quality objectives file for frequency and completeness.  The warnings are suppressed. 

```{r, warning = T}
qcMWRfre(res = resdat, frecom = frecomdat, warn = FALSE)
```

The output shows the frequency checks from the combined files.  Each row applies to a frequency check for a parameter. The `Parameter` column shows the parameter, the `obs` column shows the total records that apply to regular activity types, the `check` column shows the relevant activity type for each frequency check, the `count` column shows the number of records that apply to a check, the `standard` column shows the relevant percentage required for the quality control check from the quality control objectives file, and the `met` column shows if the standard was met by comparing if `percent` is greater than or equal to `standard`.

## Quality control for completeness

The quality control checks for completeness are used to assess the number of regular samples relative to qualified samples that apply to each parameter. Regular samples are those with `Field Msr/Obs` or `Sample-Routine` in the `Activity Type` column of the results file and qualified samples are those marked as `Q` in the `Result Measure Qualifier` column of the results file.  The completeness checks require results data (as in `resdat` above) and a data quality objectives file for frequency and completeness (as in `frecomdat`).

The `tabMWRcom()` function is used to create a table that shows the completeness percentages for each parameter.  As explained in the previous section, the function can be used with inputs as paths to the relevant files or as data frames returned by `readMWRresults()` and `read_frecom()`.  

A summary table showing the number of data records, number of qualified records, and percent completeness can be created with `tabMWRcom()`.  The `% Completeness` column shows cells as green or red if the required percentage of observations for completeness are present as specified in the data quality objectives file.  The `Hit/Miss` column shows similar information but in text format, i.e., `MISS` is shown if the quality control standard for completeness is not met. The function is also executed with the data frames from above, since they have already passed the internal checks. 

```{r, warning = T}
tabMWRcom(res = resdat, frecom = frecomdat)
```

The `tabMWRcom()` function uses the `qcMWRcom()` function internally to create the table.  This function creates the raw summaries of completeness from the input data.  Typically, `qcMWRcom()` is not used by itself, but it is explained here to demonstrate how the raw summaries can be obtained. 

Below, the `qcMWRcom()` function is executed with the data frames for the results and quality objectives file for completeness.  The warnings are suppressed. 

```{r, warning = T}
qcMWRcom(res = resdat, frecom = frecomdat, warn = FALSE)
```

The output shows the completeness checks from the combined files.  Each row applies to a completeness check for a parameter. The `datarec` and `qualrec` columns show the number of data records and qualified records, respectively. The `datarec` column specifically shows only records not for quality control by excluding those as duplicates, blanks, or spikes in the count. The `standard` column shows the relevant percentage required for the quality control check from the quality control objectives file, the `complete` column shows the calculated completeness taken from the input data, and the `met` column shows if the standard was met by comparing if `complete` is greater than or equal to `standard`.
